---
title: "CrimStats_Chicago"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install R libraries

```{r}


library(lubridate)
library(readr)
library(dplyr)
library(DT)
library(ggrepel)
library(leaflet)
li?rary(ggplot2)
library(scales)

#if(!require(lubridate)) install.packages("lubridate")
# if(!require(readr)) install.packages("readr")
# if(!require(dplyr)) install.packages("dplyr")
# if(!require(DT)) install.packages("DT")
# if(!require(ggrepel)) install.?ackages("ggrepel")
# if(!require(leaflet)) install.packages("leaflet")
# if(!require(ggplot2)) install.packages("ggplot2")
# if(!require(scales)) install.packages("scales")


```

## Data Preprocessing

Read the data
Load the data using readr and read_csv(?.

```{r}
getwd()
library(readr)
# path <- "http://spatial.binghamton.edu/projects/crime/data/SF_Crime_2007_2016.csv"
path <- "./Crimes__2001_to_Present.csv"
df <- read_csv(path)
```


## Display Data
Display the data using DT and datatable().

```{r}
libr?ry(DT)

df$Time <- format(as.POSIXlt(df$Date,format="%m/%d/%Y %I:%M:%S %p"),"%I:%M:%S %p ")

df$Date_Sub <- format(as.POSIXct(df$Date,format="%m/%d/%Y %H:%M:%S"),"%m/%d/%Y")

df_sub <- df[1:100,]  # display the first 100 rows

library(lubridate)

# class(d?_sub$Date)
# 
# ss <- as.Date(df_sub$Date)

# df_sub$Time <- format(as.POSIXlt(df_sub$Date,format="%m/%d/%Y %I:%M:%S %p"),"%I:%M:%S %p ")
# 
# df_sub$Date_Sub <- format(as.POSIXct(df_sub$Date,format="%m/%d/%Y %H:%M:%S"),"%m/%d/%Y")
# 

# df_sub$Time <- as.?haracter(df_sub$Time) 

df_sub$Time <- as.character(df_sub$Time) 
datatable(df_sub, options = list(pageLength = 5,scrollX='400px'))

# colnames(dftest)
colnames(df)

df <- df %>% select(ncol(df), (ncol(df)-1), everything())

ncol(df)
```

```{r}
sprintf("N?mber of Rows in Dataframe: %s", format(nrow(df),big.mark = ","))
head(df)
```

## Preprocess Data
The All-Caps text is difficult to read. Let's force the text in the appropriate columns into proper case.

```{r}
# str(df)
proper_case <- function(x) {
  ret?rn (gsub("\\b([A-Z])([A-Z]+)", "\\U\\1\\L\\2" , x, perl=TRUE))
}


library(stringr)
names(df) <- str_replace_all(names(df), c(" " = "_" , "," = "" ))

# df_testy = df

str(df)

df$Day <- weekdays(as.Date(df$Date_Sub))

# df <- df %>% select(-day)


library?dplyr)
df <- df %>% mutate(Primary_Type = proper_case(Primary_Type),
                    Description = proper_case(Description),
                    Location_Description = proper_case(Location_Description),
                    Arrest = proper_case(Arrest),?                    Block = proper_case(Block),
                    Time = as.character(Time))

# class(df_test)

df_sub <- df[1:100,]  # display the first 100 rows
datatable(df_sub, options = list(pageLength = 5,scrollX='400px'))


```


## Visualize Data?Crime across space
Display crime incident locations on the map using leaflet. Click icons on the map to show incident details.


```{r}
library(leaflet)

data <- df[1:10000,] # display the first 10,000 rows
data$popup <- paste("<b>Case #: </b>", data$Case_?umber, "<br>", "<b>Category: </b>", data$Primary_Type,
                    "<br>", "<b>Description: </b>", data$Description,
                    "<br>", "<b>Description: </b>", data$Location_Description,
                    "<br>", "<b>Day of week: </b>", ?ata$Day,
                    "<br>", "<b>Date: </b>", data$Date_Sub,
                    "<br>", "<b>Time: </b>", data$Time,
                    "<br>", "<b>PD Ward: </b>", data$Ward,
                    "<br>", "<b>Arrest: </b>", data$Arrest,
            ?       "<br>", "<b>Address: </b>", data$Block,
                    "<br>", "<b>Longitude: </b>", data$Longitude,
                    "<br>", "<b>Latitude: </b>", data$Latitude)

leaflet(data, width = "100%") %>% addTiles() %>%
  addTiles(group = "OSM (defa?lt)") %>%
  addProviderTiles(provider = "Esri.WorldStreetMap",group = "World StreetMap") %>%
  addProviderTiles(provider = "Esri.WorldImagery",group = "World Imagery") %>%
  # addProviderTiles(provider = "NASAGIBS.ViirsEarthAtNight2012",group = "Nighttime ?magery") %>%
  addMarkers(lng = ~Longitude, lat = ~Latitude, popup = data$popup, clusterOptions = markerClusterOptions()) %>%
  addLayersControl(
    baseGroups = c("OSM (default)","World StreetMap", "World Imagery"),
    options = layersControlOptions(col?apsed = FALSE)
  )
```

## Crime Over Time

```{r}
df_crime_daily <- df %>%
    mutate(Date = as.Date(Date_Sub, "%m/%d/%Y")) %>%
  group_by(Date) %>%
  summarize(count = n()) %>%
  arrange(Date)

library(ggplot2)
library(scales)
plot <- ggplot(df_crime_dai?y, aes(x = Date, y = count)) +
  geom_line(color = "#F2CA27", size = 0.1) +
  geom_smooth(color = "#1A1A1A") +
  # fte_theme() +
  scale_x_date(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  labs(x = "Date of Crime", y = "Number of Crimes"? title = "Daily Crimes in Chicago from 2001 - 2021")
plot
```


## Aggregate Data
Summarize the data by incident category.


```{r}
df_category <- sort(table(df$Primary_Type),decreasing = TRUE)
df_category <- data.frame(df_category[df_category > 10000])
co?names(df_category) <- c("Category", "Frequency")
df_category$Percentage <- df_category$Frequency / sum(df_category$Frequency)
datatable(df_category, options = list(scrollX='400px'))

```


Create a bar plot based on the incident category.


```{r}
library(?gplot2)
library(ggrepel)
bp<-ggplot(df_category, aes(x=Category, y=Frequency, fill=Category)) + geom_bar(stat="identity") + 
  theme(axis.text.x=element_blank()) + geom_text_repel(data=df_category, aes(label=Category))
bp
```



```{r}
bp<-ggplot(df_catego?y, aes(x="", y=Percentage, fill=Category)) + geom_bar(stat="identity") 
pie <- bp + coord_polar("y") 
pie
```



```{r}
install.packages("sf")
library(sf)

getwd()

chi_neighborhoods <- st_read("./Neighborhoods_2012/Neighborhoods_2012b.shp")

```

```{r}
p?ot(df$geometry)

st_crs(chi_neighborhoods)

df <- df %>% mutate(geometry = df$Location)

df_sorted <- df %>% arrange(df$Date_Sub)

head(df_sorted)

##Remove Rows with NAs

df_nonas <-  df[complete.cases(df), ]

df_nonas <- df_nonas %>% arrange(df_nonas$Dat?_Sub)

head(df_nonas)

chi_df <- st_as_sf(df_nonas, coords = c("Longitude", "Latitude"), crs = "+proj=longlat +ellps=WGS84 +no_defs")

chi_df <- st_transform(chi_df, crs = st_crs(chi_neighborhoods))

head(chi_df)

crime_agg <- st_join(chi_df, chi_neighborh?ods)

head(crime_agg)


```

Aggregate Crimes by area
```{r}

crime_agg$number_crimes <- 1

head(crime_agg)
head(crime_agg)
crime_agg[crime_agg$Primary_Type == "Theft"]

crime_agg_theft <- filter(crime_agg, Primary_Type == "Theft")



crime_types <- unique?crime_agg$Primary_Type)

#Theft aggregate Chicago
aggregate_theft <- aggregate(number_crimes ~ PRI_NEIGH, data = crime_agg_theft, FUN = sum)

summary(aggregate_theft$number_crimes)

nrow(aggregate_theft)

nrow(chi_neighborhoods)

```


```{r}

#Create freq?ency table for crimes by neighborhood
crime_agg_category <- table(crime_agg$PRI_NEIGH,crime_agg$Primary_Type)


#Convert Table to DF
crime_agg_category_df <- as.data.frame.matrix(crime_agg_category) 

crime_agg_category_df$PRI_NEIGH = neigh_names

crime_ag?_category_df <- crime_agg_category_df %>% select(ncol(crime_agg_category_df), everything())

rownames(crime_agg_category_df) <-  seq(length=nrow(crime_agg_category_df))


#crime_agg_category_year <- table(crime_agg$PRI_NEIGH,crime_agg$Primary_Type, crime_a?g$Year)


chi_neighborhoods_crimes <- left_join(chi_neighborhoods, crime_agg_category_df)

head(chi_neighborhoods)


crime_agg_category <- data.frame(crime_agg_category[crime_agg_category > 10000])
colnames(crime_agg_category) <- c("Category", "Frequency")?crime_agg_category$Percentage <- crime_agg_category$Frequency / sum(crime_agg_category$Frequency)
datatable(crime_agg_category, options = list(scrollX='400px'))
```

## Making choropleth maps
Finally we are ready to make some choropleth maps.

For these ma?s we are going to use ggplot2 again so we need to load it.


```{r}
library(ggplot2)
ggplot(chi_neighborhoods_crimes, aes(fill = chi_neighborhoods_crimes$Arson)) +
  geom_sf() +  scale_fill_gradient(low = "white", high = "red") + 
  labs(fill = "# of arson?",
       title = "Arsons in Chicago, by neighborhood",
       subtitle = "2001 - 2021")  +  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank())

```

```{r}

getwd()
save(chi_neighborhoods_crim?s, file = "../chi_neighborhoods_crimes.rda")

```



## Interactive Maps


```{r}

## Transform Data to right projection since leaflet expects standard CRS, WGS84 using long and lat -- projected means no long lat
chi_neighborhoods_crimes <- st_transform(ch?_neighborhoods_crimes,
                                        crs = "+proj=longlat +datum=WGS84")

pal <- colorNumeric("OrRd", chi_neighborhoods_crimes$`Human Trafficking`)

leaflet()  %>% 
  addTiles('http://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png', 
?          attribution = '&copy; <a href="http://openstreetmap.org">
                OpenStreetMap</a> contributors') %>%
   addPolygons(data = chi_neighborhoods_crimes$geometry,
              color = "black",
              weight = 1,
               popup ? paste0("Neighborhood: ", chi_neighborhoods_crimes$PRI_NEIGH,
                             "<br>",
                             "Number of Humans Trafficked: ", chi_neighborhoods_crimes$`Human Trafficking`),fillColor = pal(chi_neighborhoods_crimes$`Human T?afficking`),
              fillOpacity = .75) %>%  addLegend(pal = pal, 
            values = chi_neighborhoods_crimes$`Human Trafficking`,
            opacity = 1, 
            title = "Number of Humans Trafficked")



```

